---
title: "BDA - Assignment 1"
author: "Anonymous"
output:
  html_document:
    toc: yes
    toc_depth: '1'
    df_print: paged
  html_notebook:
    toc: yes
    toc_depth: '1'
    df_print: paged
---

# Questions

See here: https://avehtari.github.io/BDA_course_Aalto/assignments/assignment1.pdf

# Question 2

## a)

```{r}
# the mean of the beta distribution
m <- 0.2

# the variance of the beta distribution
v <- 0.001
```


```{r}
alpha <- m * ((m * (1 - m) / v) - 1)
beta <- alpha * (1 - m) / m
```

```{r}
x <- seq(0, 1, 1/100)
y <- dbeta(x, alpha, beta)

plot(x, y)
```

## b)

```{r}
X <- rbeta(1000, alpha, beta)
hist(X, freq = FALSE)
```

## c)

```{r}
sample_mean <- mean(X)
sample_var <- var(X)

cat(abs(sample_mean - m), "\n")
cat(abs(sample_var == v), "\n")
```


## d)

```{r}
quantile(X, c(5/100, 95/100))
```
# Question 3

Extracted probabilities from the problem:

```{r}
prob_pos_given_lungcancer <- 98/100
prob_neg_given_nolungcancer <- 96/100
prob_lungcancer <- 1/1000
```
Verify the probabilities from the given hints:

```{r}
prob_pos_given_nolungcancer <- 1 - prob_neg_given_nolungcancer
abs(prob_pos_given_nolungcancer - 4/100) < 1/1000
```

```{r}
prob_pos_and_lungcancer <- prob_pos_given_lungcancer * prob_lungcancer

abs(prob_pos_and_lungcancer - 0.098/100) < 1/1000
```
Calculate probability of false negative i.e probability negative given cancer:

```{r}
prob_false_neg <- 1 - prob_pos_given_lungcancer
prob_false_neg
```

Calculate probability of false positive i.e. probability positive given no cancer:

```{r}
prob_false_pos <- prob_pos_given_nolungcancer
prob_false_pos
```
Probability have cancer given negative result:

```{r}
prob_nolungcancer <- 1 - prob_lungcancer 
prob_pos_and_nolungcancer <- prob_pos_given_nolungcancer * prob_nolungcancer

prob_pos <- prob_pos_and_lungcancer + prob_pos_and_nolungcancer
prob_neg <- 1 - prob_pos

prob_neg_given_lungcancer <- 1 - prob_pos_given_lungcancer
prob_neg_and_lungcancer <- prob_neg_given_lungcancer * prob_lungcancer

prob_lungcancer_given_neg <- prob_neg_and_lungcancer/prob_neg
prob_lungcancer_given_neg
```
Therefore, probability lung cancer does not get detected is almost 0.

Probability have cancer given positive result:

```{r}
prob_lungcancer_given_pos <- prob_pos_and_lungcancer/prob_pos
prob_lungcancer_given_pos
```
Therefore, probability un-necessarily administer medication is about 2.3%.

# Question 4


```{r}
p_red <- function(boxes){
  prob_A <- 0.4
  prob_B <- 0.1
  prob_C <- 1 - (prob_A + prob_B)

  box_totals <- apply(boxes, 1, sum)
  
  prob_red_and_A <- boxes["A", "red"]/box_totals["A"] * prob_A
  prob_red_and_B <- boxes["B", "red"]/box_totals["B"] * prob_B
  prob_red_and_C <- boxes["C", "red"]/box_totals["C"] * prob_C

  prob_red_and_A + prob_red_and_B + prob_red_and_C
}
```

```{r}
p_box <- function(boxes) {
  prob_A <- 0.4
  prob_B <- 0.1
  prob_C <- 1 - (prob_A + prob_B)

  box_totals <- apply(boxes, 1, sum)

  prob_red_and_A <- boxes["A", "red"]/box_totals["A"] * prob_A
  prob_red_and_B <- boxes["B", "red"]/box_totals["B"] * prob_B
  prob_red_and_C <- boxes["C", "red"]/box_totals["C"] * prob_C

  prob_red <- prob_red_and_A + prob_red_and_B + prob_red_and_C

  prob_A_given_red <- prob_red_and_A / prob_red
  prob_B_given_red  <- prob_red_and_B / prob_red
  prob_C_given_red <- prob_red_and_C / prob_red

  c(prob_A_given_red, prob_B_given_red, prob_C_given_red)
}
```
# Question 5

From the problem:

```{r}
# can be different sex
prob_frat <- 1/150

# must be same sex
prob_ident <- 1/400

prob_boy <- 1/2
prob_girl <- 1 - prob_boy
```

Elvis had a twin brother.

Find probability Elvis is an identical twin given sibling is a boy:

```{r}
p_identical_twin <- function(fraternal_prob, identical_prob){
  # Note: you == Elvis
  prob_you_identical <- identical_prob
  prob_you_fraternal <- fraternal_prob

  prob_boy_sibling_given_you_identical <- 1
  prob_boy_sibling_given_you_fraternal <- prob_boy

  prob_you_identical_and_boy_sibling <- prob_boy_sibling_given_you_identical * prob_you_identical
  
  prob_boy_sibling_and_you_identical <- prob_you_identical_and_boy_sibling
  prob_boy_sibling_and_you_fraternal <- prob_boy_sibling_given_you_fraternal * prob_you_fraternal

  prob_boy_sibling <- prob_boy_sibling_and_you_identical + prob_boy_sibling_and_you_fraternal

  prob_you_identical_given_boy_sibling <- prob_you_identical_and_boy_sibling/prob_boy_sibling
  
  return(prob_you_identical_given_boy_sibling)
}
```

# Check Answer

```{r}
library(markmyassignment)

assignment_path <-
  paste(
    "https://github.com/avehtari/BDA_course_Aalto/",
    "blob/master/assignments/tests/assignment1.yml",
    sep = ""
  )
set_assignment(assignment_path)
mark_my_assignment()

```

